---
title: "Package: NAMCS"
author: "Alan Brookhart"
date: "1/22/2017"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NAMCS NSAID data

This package builds a cohort from the 2005-2009 public use files from National Ambulatory Medical Care Data Survey (NAMCS).  The NAMCS survey covers over 20,000 visits per year and includes data on patient demographics, comorbidities, physician and practice characteristics, and treatment received, including medications. Currently, medications are classified using the Multum Lexicon. Each year a few variables are added to, deleted from, or recoded in the public use dataset. 


From these data, we identify new users of either NSAID or Cox-2 selective non-steroidal antiinflammatory drugs (NSAIDs). The data are augmented with a simulated peptic ulcer disease outcome. Variables were selected for this extract based on relevance to the example analyses and on availability for most or all of the 2005-2009 time range.

## Installation

You can install the package from Github

```{r bb-installation,eval=FALSE}
library("devtools")
devtools::install_github("alanbrookhart/NAMCS")
```

### Variable names

```{r data}
library("NAMCS")
names(ns)
```


## Codebook 

* age = age in years of participant (years)
* sex = sex of participant (Male vs Female)
* year = year of the survey
* race = paricipant race (Hispanic, White, Black, Other)
* region = region of the country (south, north east, mid-west, west)
* arthritis = history of osteoarthritis
* asthma = history of asthma 
* cancer = history of cancer
* cerebrovascular_disease = history of cerebrovascular disease
* chronic_kidney_disease = history of chronic renal disease
* heart_failure = history of heart failure 
* chronic_pulmonary_disease = history of chronic obstructive pulmonary disease
* depression = history of depression
* diabetes = history of diabetes
* hyperlipidemia = history of hyperlipidemia
* mi = history of myocardial infarction
* hypertension = history of hypertension
* coronary_artery_disease = history of ischemic heart disease
* osteoporosis = history of osteoporosis
* obesity = patient is obese 
* anti_hypertensive_use = history of use of antihypertensives
* h2_antagonist_use = history of use of H2 blockers, gastroprotective agents
* ppi_use = history of use of proton pump inhibitors, gastroprotective agents
* aspirin_use = history of use of aspirin
* anti_coagulant_use = history of use of anticoagulants
* corticosteroid_use = history of use of steroids
* cox2_initiation = whether the patient initiated a coxib vs a NSAID
* incident_pud = new diagnosis peptic ulcer disease durig follow-up (simulated


### Summary 

```{r data2}
summary(ns) 
```


## NAMCS Statin cohort data

We have built a cohort from the 2005-2009 public use files from National Ambulatory Medical Care Data Survey (NAMCS).  The NAMCS survey covers over 20,000 visits per year and includes data on patient demographics, comorbidities, physician and practice characteristics, and treatment received, including medications. Currently, medications are classified using the Multum Lexicon. Each year a few variables are added to, deleted from, or recoded in the public use dataset. 

From these data, we identify all visits among adults who are not currently treated with a statin.  Patients are then either observed to initiate a statin or remain untreated.  Statin initiators are further classified into users of high vs low potency statins. 

We then use the baseline data to simulate two outcomes: risk of a cardiovascular disease event and death.  We further simulate the time to protocol non-adherence (initation of a statin among patients who are untreated or discontinuation of a statin among initiators.)


## Notes

NAMCS sampling weights are not included.


