---
title: "National Ambulatory Medical Care Survey, Statin study"
author: "Alan Brookhart"
date: "6/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(foreign)
library(causalRisk)
library(foreach)
library(purrr)
library(NAMCS)
```

# Exercise 1

## Initial exploration of data

### Unweighted Table 1

```{r table1}

models = specify_models(identify_outcome(death_time),
                        identify_treatment(statin_use))

death_unadj = estimate_ipwrisk(sta, models, labels  = c("Death, unadjusted"))

make_table1(death_unadj,
            sex, 
            race, 
            region, 
            age, obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use, smd = TRUE)
```


## Estimation of simple propensity score with just age, sex, and race

### PS Histogram
```{r hist}

models = specify_models(identify_outcome(death_time),
                        identify_treatment(statin_use, ~age+ sex +race))

death_agesex = estimate_ipwrisk(sta, models, labels  = c("Death, age-sex adjusted"))


hist(death_agesex)
```

### IP weighted Table 1

```{r ipw_table1}
make_table1(death_agesex,
            sex, 
            race, 
            region, 
            age, obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use, smd = TRUE)
```


### IP weights summary table
```{r wt_summary}
make_wt_summary_table(death_agesex)

```

## Estimation of elaborate propensity score 

### PS Histogram
```{r hist2}
models = specify_models(identify_outcome(death_time),
                        identify_treatment(statin_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

death_bigps = estimate_ipwrisk(sta, models, labels  = c("Death, fully adjusted"))

hist(death_bigps)
```

### IP weighted table 1

```{r ipw2_table1}
make_table1(death_bigps,
            sex, 
            race, 
            region, 
            age, 
            obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use, smd = TRUE)
```


### IP weights summary table
```{r wt_summary2}
make_wt_summary_table(death_bigps)

```

## SMR weighting

### Unweighted PS Histogram

```{r hist_smr}
models = specify_models(identify_outcome(death_time),
                        identify_treatment(statin_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

death_bigps_smr1 = estimate_ipwrisk(sta, models, wt_type = 1, labels  = c("Death, adjusted, SMRW Untreated"))

death_bigps_smr2 = estimate_ipwrisk(sta, models, wt_type = 2, labels  = c("Death, adjusted, SMRW Treated"))

hist(death_bigps)
```

### Weighted PS histograms
```{r}
hist(death_bigps, death_bigps_smr1, death_bigps_smr2, ncol = 3, weight = TRUE)
```

### Table 1 with SMR weights standardizing the untreated
```{r}
make_table1(death_bigps_smr1,
            sex, 
            race, 
            region, 
            age, obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use, side.by.side = T, smd = TRUE)
```

### Table 1 with SMR weights standardizing the treated
```{r}
make_table1(death_bigps_smr2,
            sex, 
            race, 
            region, 
            age, obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use,  side.by.side = T, smd = TRUE)
```

## Estimation of elaborate propensity score w/ trimming at top and bottom 5%

### PS Histogram
```{r hist3}
models = specify_models(identify_outcome(death_time),
                        identify_treatment(statin_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

death_bigps_trim = estimate_ipwrisk(sta, models, labels  = c("Death, adjusted, trimmed"), trim = 0.05)

hist(death_bigps_trim)
```

### IP weighted table 1

```{r ipw3_table1}
make_table1(death_bigps_trim,
            sex, 
            race, 
            region, 
            age, obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use, smd = TRUE)
```

### IP weight summary table
```{r wt_summary3}
make_wt_summary_table(death_bigps_trim)
```

### Effect of removing hyperlipidemia from the PS model
```{r remove_hyperlipid}

death_bigps2 = death_bigps %>%
  update_treatment(new_formula = ~. -hyperlipidemia) %>%
  update_label("Drop Hyperlipidemia") %>%
  re_estimate()

hist(death_bigps, death_bigps2)

make_table1(death_bigps2,
            sex, 
            race, 
            region, 
            age, obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use, smd = TRUE)

```


# Exercise 2

## Estimation of outcome risk

### Overall risk death, unadjusted
```{r overall}
models = specify_models(identify_outcome(death_time))

death_overall = estimate_ipwrisk(sta, models, labels  = c("Death overall"))

plot(death_overall) + ylab("Cumulative Risk") + xlab("Time in years")
```


## Overall risk CV events, unadjusted, censoring by death
```{r cv_overall}
models = specify_models(identify_outcome(cv_time),
                        identify_censoring(death_time))

cv_overall = estimate_ipwrisk(sta, models, labels  = c("CV risk overall"))

plot(cv_overall) + ylab("Cumulative Risk") + xlab("Time in years")
```

## Estimation of risk death, by treatment group

### Unadjusted cumulative risk of death and unadjusted cumulative risk difference

```{r}
plot(death_unadj) + 
  ylab("Cumulative Risk") + xlab("Time in years")

plot(death_unadj, effect_measure_type = "RD") +
  ylab("Cumulative Risk Difference") + xlab("Time in years")
```

### Adjusted and unadjusted cumulative risk of death and unadjusted cumulative risk difference

```{r}
plot(death_unadj, death_bigps,scales = "fixed") + 
  ylab("Cumulative Risk") + xlab("Time in years")

plot(death_unadj, death_bigps, effect_measure_type = "RD", scales = "fixed") +
  ylab("Cumulative Risk Difference") + xlab("Time in years")
```


### Adjusted, unadjusted, trimmed, and SMR weighted cumulative risk of death and unadjusted cumulative risk difference

```{r}
plot(death_unadj, death_agesex, death_bigps, death_bigps_trim, death_bigps_smr1, death_bigps_smr2, ncol = 3, scales = "fixed") + 
  ylab("Cumulative Risk") + xlab("Time in years")

plot(death_unadj, death_agesex, death_bigps, death_bigps_trim, death_bigps_smr1, death_bigps_smr2, ncol = 3, effect_measure_type = "RD", scales = "fixed") +
  ylab("Cumulative Risk Difference") + xlab("Time in years")

```


### Table 2: 10-year cumulative risk and risk difference

```{r}
make_table2(death_unadj, death_agesex, death_bigps, death_bigps_trim, effect_measure_type = "RD", risk_time = 10) 

forest_plot(death_unadj, death_agesex, death_bigps, death_bigps_trim, death_bigps_smr1, death_bigps_smr2, risk_time = 10, effect_measure_type = "RD")

```

### Table 2: 10-year cumulative risk and risk ratio

```{r}
make_table2(death_unadj, death_agesex, death_bigps, death_bigps_trim, effect_measure_type = "RR", risk_time = 10) 
```

## Estimation of the risk a CV event

### Estimation w/ death as censoring event

```{r cv_unadj}
models = specify_models(identify_outcome(cv_time),
                        identify_censoring(death_time),
                        identify_treatment(statin_use))

cv_unadj = estimate_ipwrisk(sta, models, labels  = c("CV Risk, unadjusted"))

plot(cv_unadj)  +
  ylab("Cumulative Risk Difference") + xlab("Time in years")
```

### Estimation w/ death as censoring event, addressing confounding 

```{r cv_iptw_adj}
models = specify_models(identify_outcome(cv_time),
                        identify_censoring(death_time),
                        identify_treatment(statin_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

cv_iptw_adj = estimate_ipwrisk(sta, models, labels  = c("CV Risk, censoring at death, IPTW"))

plot(cv_unadj, cv_iptw_adj, scales = "fixed")  +
  ylab("Cumulative Risk Difference") + xlab("Time in years")
```

### Estimation w/ death as censoring event, addressing confounding and dependent censoring

```{r cv_iptcw_adj}

models = specify_models(identify_outcome(cv_time),
                        identify_censoring(death_time, ~age +
                                          sex +
                                          race +
                                          region + 
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use),
                        identify_treatment(statin_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

cv_iptcw_adj = estimate_ipwrisk(sta, models, labels  = c("CV Risk, cens at death, IPTCW"))

plot(cv_unadj, cv_iptw_adj, cv_iptcw_adj, scales = "fixed", ncol = 3)  + 
  ylab("Cumulative Risk") + xlab("Time in years")

plot(cv_unadj, cv_iptw_adj, cv_iptcw_adj, effect_measure_type = "RD", overlay = TRUE,  ncol = 3, scales = "fixed")  +
  ylab("Cumulative Risk Difference") + xlab("Time in years")
```

### Estimation w/ death as a competing event

```{r cv_cr_adj}
models = specify_models(identify_outcome(cv_death_time),
                        identify_competing_risk(cv_indicator, event_value = 1),
                        identify_treatment(statin_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

cv_cr_adj = estimate_ipwrisk(sta, models, labels  = c("CV Risk, death as comp risk, IPTW"))

plot(cv_iptcw_adj, cv_cr_adj, ncol = 2, scales = "fixed")  +
  ylab("Cumulative Risk Difference") + xlab("Time in years")

plot(cv_iptcw_adj, cv_cr_adj, effect_measure_type = "RD", ncol = 2, scales = "fixed")  +
  ylab("Cumulative Risk Difference") + xlab("Time in years")

make_table2(cv_iptcw_adj, cv_cr_adj, risk_time = 10)
```


# Exercise 3

## Consideration of >2 treatment groups, analysis by statin potency

### Unweighted Table 1

```{r table1_multi}
models = specify_models(identify_outcome(death_time),
                        identify_treatment(statinpotency_use))

death_unadj_multi = estimate_ipwrisk(sta, models, labels  = c("Death, by potency, unadjusted"))

make_table1(death_unadj_multi,
            sex, 
            race, 
            region, 
            age, obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use, smd = TRUE)
```


## Estimation w/ an elaborate propensity score 

### PS Histogram
```{r hist2_multi}
models = specify_models(identify_outcome(death_time),
                        identify_treatment(statinpotency_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

death_bigps_multi = estimate_ipwrisk(sta, models, labels  = c("Death, by potency, adjusted"))

hist(death_bigps_multi, cat = 1, binwidth = 0.03)
hist(death_bigps_multi, cat = 2, binwidth = 0.03)
```

### IP weighted table 1

```{r ipw2_table1_multi}
make_table1(death_bigps_multi,
            sex, 
            race, 
            region, 
            age, obesity,
            diabetes, 
            tobacco_use,
            asthma, 
            hyperlipidemia, 
            heart_failure, 
            cerebrovascular_disease, 
            chronic_kidney_disease, 
            chronic_pulmonary_disease, 
            osteoporosis, 
            cancer, 
            ppi_use, 
            aspirin_use, 
            anti_coagulant_use, 
            corticosteroid_use, smd = TRUE)
```


### IP weight summary table
```{r wt_summary_multi}
make_wt_summary_table(death_bigps_multi)

```

### Estimate of cumulative risk of CV outcomes and death by statin potency 
```{r risk_multi}
models = specify_models(identify_outcome(cv_time),
                        identify_censoring(death_time),
                        identify_treatment(statinpotency_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

cv_bigps_multi = estimate_ipwrisk(sta, models, labels  = c("CV Risk, by potency"))

plot(death_bigps_multi , cv_bigps_multi, ncol = 2, scales = "fixed")  +
  ylab("Cumulative Risk") + xlab("Time in years")

plot(death_bigps_multi , cv_bigps_multi, effect_measure_type = "RD", ncol = 2, scales = "fixed")  +
  ylab("Cumulative Risk Difference") + xlab("Time in years")

make_table2(death_bigps_multi , cv_bigps_multi, risk_time = 10)
```


## Estimation of risk if nonadherence could be prevented (per protocol effect)

## Non-adherence risk by treatment group, censoring by death
```{r nonadh_overall}
models = specify_models(identify_outcome(nonadherence),
                        identify_treatment(statin_use),
                        identify_censoring(death_time))

nonadherence_overall = estimate_ipwrisk(sta, models, labels  = c("Nonadherence risk, unadjusted"))

plot(nonadherence_overall) + ylab("Cumulative Risk") + xlab("Time in years")
```


### Effect of uninterrupted statin treatment on risk death
```{r pp_death}
models = specify_models(identify_outcome(death_time),
                        identify_censoring(nonadherence),
                        identify_treatment(statin_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

death_bigps_pp = estimate_ipwrisk(sta, models, labels  = c("Death, per protocol"))

plot(death_bigps, death_bigps_pp, scales = "fixed")   +
  ylab("Cumulative Risk") + xlab("Time in years")

plot(death_bigps, death_bigps_pp, effect_measure_type = "RD", scales = "fixed")   +
  ylab("Cumulative Risk") + xlab("Time in years")

make_table2(death_bigps, death_bigps_pp, risk_time = 10)
```


### Effect of uninterrupted statin treatment on risk of cardiovascular events
```{r pp_dcv}
models = specify_models(identify_outcome(cv_time),
                        identify_censoring(nonadherence),
                        identify_censoring(death_time),
                        identify_treatment(statin_use, ~age +
                                          sex +
                                          race + 
                                          region +
                                          obesity +
                                          diabetes+ 
                                          tobacco_use+
                                          asthma+ 
                                          hyperlipidemia+ 
                                          heart_failure+ 
                                          cerebrovascular_disease+ 
                                          chronic_kidney_disease+ 
                                          chronic_pulmonary_disease+ 
                                          osteoporosis+ 
                                          cancer+ 
                                          ppi_use+ 
                                          aspirin_use+ 
                                          anti_coagulant_use+ 
                                          corticosteroid_use))

cv_bigps_pp = estimate_ipwrisk(sta, models, labels  = c("CV, per protocol"))

plot(cv_iptw_adj, cv_bigps_pp, scales = "fixed")   +
  ylab("Cumulative Risk") + xlab("Time in years")

plot(cv_iptw_adj, cv_bigps_pp, effect_measure_type = "RD", scales = "fixed")   +
  ylab("Cumulative Risk") + xlab("Time in years")

make_table2(cv_iptcw_adj, cv_bigps_pp, risk_time = 10)
```

